---
title: "Inmates in Catalonia and Spain"
author: "pacha"
date: "November 8, 2017"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE, tidy = FALSE}
knitr::opts_chunk$set(message = F, error = F, warning = F, comment = NA, fig.align = 'center', dpi = 100, fig.width=9, fig.height=8, tidy = F, cache = T, dev = 'CairoPNG')
```

# WORK IN PROGRESS

Run `install.packages("prettydoc")` just once before composing this document. Without prettydoc it won't work.

## 1st: Import data

Tables downloaded from [Idescat](https://www.idescat.cat/economia/inec?tc=3&id=de03) in Excel format. 

Each dataset has 6 columns: 

* Year
* Number of convicted inmates
* Percentage of convicted inmates from the total
* Number of preventive inmates
* Percentage of preventive inmates from the total
* Total number of inmates

*This dataset is completely messy!*

```{r download_and_tidy}
if (!require("pacman")) install.packages("pacman")
p_load(readr,tidyr,dplyr,stringr,readxl,ggplot2,viridis)
p_load_gh("ropenscilabs/skimr")

url = "https://www.idescat.cat/economia/inec?tc=3&id=de03&dt=201600&ex=1"
file = "immates_data.xls"
if(!file.exists(file))(try(download.file(url,file)))

################################

# Catalunya

immates_catalunya = read_excel("immates_data.xls", sheet = "Full 2", col_names = FALSE, na = "..", skip = 7, n_max = 4) %>% 
  setNames(c("category",paste("value", 2014:2016, sep = "_"), paste("percent_change", 2014:2016, sep = "_"))) %>% 
  filter(category != "Penats (%)", category != "Preventius (%)")

immates_catalunya_values = immates_catalunya %>% select(category, matches("value")) %>% 
  gather(year, value, -category) %>% 
  mutate(country = "Catalunya",
         year = as.integer(str_replace(year, "value_", "")),
         category = if_else(category == "Penats (nombre)", "convicted", "preventive imprisonment")) %>% 
  select(country, category, year, value)
  
immates_catalunya_percents = immates_catalunya %>% select(category, matches("percent_change")) %>% 
  gather(year, percent_change, -category) %>% 
  mutate(country = "Catalunya",
         year = as.integer(str_replace(year, "percent_change_", "")),
         category = if_else(category == "Penats (nombre)", "convicted", "preventive imprisonment")) %>% 
  select(country, category, year, percent_change)

rm(immates_catalunya)

################################

# Spain

immates_spain_values = read_excel("immates_data.xls", sheet = "Full 3", col_names = FALSE, na = ":", skip = 7, n_max = 4) %>% 
  setNames(c("category",paste("value", 2014:2016, sep = "_"), paste("percent_change", 2014:2016, sep = "_"))) %>% 
  filter(category != "Penats (%)", category != "Preventius (%)") %>% 
  select(category, matches("value")) %>% 
  gather(year, value, -category) %>% 
  mutate(country = "Spain",
         year = as.integer(str_replace(year, "value_", "")),
         category = if_else(category == "Penats (nombre)", "convicted", "preventive imprisonment")) %>% 
  select(country, category, year, value)
  
immates_spain_percents = read_excel("immates_data.xls", sheet = "Full 3", col_names = FALSE, na = "..", skip = 7, n_max = 4) %>% 
  setNames(c("category",paste("value", 2014:2016, sep = "_"), paste("percent_change", 2014:2016, sep = "_"))) %>% 
  filter(category != "Penats (%)", category != "Preventius (%)") %>% 
  select(category, matches("percent_change")) %>% 
  gather(year, percent_change, -category) %>% 
  mutate(country = "Spain",
         year = as.integer(str_replace(year, "percent_change_", "")),
         category = if_else(category == "Penats (nombre)", "convicted", "preventive imprisonment")) %>% 
  select(country, category, year, percent_change)

################################

# Tidy Data
immates_values = bind_rows(immates_catalunya_values, immates_spain_values)
immates_percents = bind_rows(immates_catalunya_percents, immates_spain_percents)

rm(immates_catalunya_percents, immates_catalunya_values)
rm(immates_spain_percents, immates_spain_values)

immates_tidy = left_join(immates_values, immates_percents, by = c("country","category", "year"))
rm(immates_values, immates_percents)

immates_tidy
```

## 2nd: Generate main statistics

Instead of classic `summary` here I'm using [Ropenscilab](https://github.com/ropenscilabs/skimr)'s `skimr`.

```{r stats}
skim(immates_tidy)
skim(immates_tidy %>% filter(country == "Catalunya"))
skim(immates_tidy %>% filter(country == "Spain"))
```

Which is the maximum of convicted inmates in Catalan prisons in a year?

```{r stats2}
immates_tidy %>% filter(value == max(value, na.rm = T))

# the same in this case
immates_tidy %>% filter(country == "Catalunya") %>%  filter(value == max(value))
immates_tidy %>% filter(country == "Catalunya") %>%  filter(value == max(value, na.rm = T))

immates_tidy %>% filter(country == "Spain") %>%  filter(value == max(value, na.rm = T))
```

## 3rd: Create a chart


```{r barplot}
ggplot(data = immates_tidy %>% mutate(label = paste0(country, " (", category, ")"))) +
  geom_bar(mapping = aes(x = year, y = value, fill = label), stat = "identity") +
  ggtitle("Immates in Catalonia and Spain") +
  theme_minimal(base_size = 13) +
  scale_fill_viridis(end = 0.85, discrete = TRUE)

ggplot(data = immates_tidy %>% mutate(label = paste0(country, " (", category, ")"))) +
  geom_bar(mapping = aes(x = year, y = value, fill = label), stat = "identity", position = "dodge") +
  ggtitle("Immates in Catalonia and Spain") +
  theme_minimal(base_size = 13) +
  scale_fill_viridis(end = 0.85, discrete = TRUE)
```

